Suplementary Material 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data

### 2023-04-18 S4_draw_IBI_predictions_EE_not_last.Rmd
remve uneccesary stuff explain

The aim of this document is to display the R code needed the reproduce the original paper figure 4 equivalent on the Estonian dataset where the last births have been ommited.

**NB!** Some of the code e.g reading in the data is same as in S1.

```{r}
#cleanup memory
gcstuff <- gc(verbose=FALSE); rm(gcstuff);

@@

```{r}
#get last birth adding function
source("./R/last_birth.R")

#simplified twinR summary tables 
source("./R/twinR_summary.R")

#fix twinR compute predictions to do prediction with no lambda  as well
source("./R/twinR_predictions.R")

#simple convenience functions
source("./R/utils.R")
```

```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- 7

#it seems that simulating the base model for Estonian not last birth data
#requires about 15GB of RAM per CPU core (tested with 61 CPU cores)

## Set option in spaMM:
spaMM::spaMM.options(nb_cores = nb_cores)
```

## Data Import Estonia

The Estonian dataset has been formatted to include the same columns as *the data_births_all* dataset from the **twinR** package. The only difference is that the columns *pop* and *monthly* are excluded as these are constant.

```{r}
#Import and preproccess Estonian Data

data_births_monthly_EE <- readRDS("./data/data_births_all_EE.rds")

#the twinR package expects population to be present
data_births_monthly_EE$pop <- "Estonia"

data_births_monthly_EE <- add_last_birth(data_births_monthly_EE)

data_births_monthly_EE_not_last <- data_births_monthly_EE[!data_births_monthly_EE$last,]

@@

@@

>>>>

The aim of this document is to display the R code needed the reproduce the original paper figure 4 equivalent on the Estonian dataset where the last births have been omitted.

@@

@@

As running the *fit_life_histories* function with scenario “**PISH**” will fit all three life history models: *fit_PP*, *fit_IBI*, and *fit_twinning.binary*. The polynomial orders for each of these models will be estimated as the one leading to the best fit between 0 and 6. The *fit_PP* and *fit_IBI* functions will include the variable twin as a predictor. The *fit_twinning.binary* function will include *maternal_id* as a random effect predictor. So, it is equivalent to running each of these three functions with their default settings and the specified scenario. Hence, to produce the figure 4 equivalent we load these fits and get the fitted models from there.

@@

Next we plot the models exactly as in the original manuscript. Note that the actual parity is +1 than the number on the graphs (because the last birth is omitted in this data, so the *parity* axis shows the parity up until the one to last).
